---
title: "Surveillance Report"
author: "Your name here"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: "cerulean"
    toc: true
params:
  district: Mountain Rural
  quantaudience: TRUE
  report_date: !r lubridate::ymd("2014-12-01")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE )
```

```{r load_packages, echo=FALSE}
# Load packages
pacman::p_load(
  rio,          # for importing data
  here,         # for locating files
  janitor,      # for data cleaning  
  lubridate,    # for date cleaning 
  flextable,    # make HTML tables
  epikit,       # formatting counts
  officer,      # table borders
  scales,       # for percents and plot dates
  gtsummary,    # for nice tables
  tidyverse     # for data management and visualization
)
```

# Summary {.tabset}

This is a demo situation report on a hypothetical outbreak of Ebola in Sierra Leone from 2014. It uses simulated data.  

## About this report  

This report was produced on `r Sys.Date()`. It uses cases reported from **`r params$district`** district, as of **`r params$report_date`**.

## Data import and cleaning

Below is the code used to import the case linelist and restrict it to relevant districts/dates.  

```{r read_data}
# Import data
surv <- import(here("data", "clean", "linelist_combined_20141201.rds")) %>% 
  filter(district == params$district) %>% 
  filter(date_report <= params$report_date | is.na(date_report))
```

## Key numbers  

```{r}
# define the number of hospitals affected
num_hosps <- surv %>%
  pull(hospital) %>%
  unique() %>%
  # na.omit() %>%  # remove NA from list, if desired
  length()

# names of the hospitals
hosp_names <- surv %>% 
  pull(hospital) %>% 
  unique() %>% 
  fct_na_value_to_level(level = "Unknown") %>% # covert NA to "Unknown"
  fct_infreq()                                 # Order by frequency
```

Key points about this outbreak:  

* In total, there have been `r nrow(surv)` cases reported, with `r epikit::fmt_count(surv, outcome == "Death")` deaths.  
* The first case was reported on `r min(surv$date_report, na.rm=T)`.   
* The last case was reported on `r max(surv$date_report, na.rm=T)`.  
* There are `r num_hosps` hospitals that have reported cases: `r paste(hosp_names, collapse = ", ")`.  




# View case linelist  

```{r show_data, eval=params$quantaudience}
surv %>% 
  head(20) %>%      # keep only the first 20 rows
  DT::datatable(
    rownames = FALSE,
    filter = "top",
    options = list(pageLength = 5, scrollX=T),
    caption = "Table 1. Surveillance line list")
```

# Case summary by hospital

```{r}
# create a summary table of cases and their outcomes by hospital
hosp_table <- surv %>% 
  drop_na(hospital) %>%       # drop cases where hospital is not known or missing
  count(hospital, outcome,    # count cases by hospital and outcome
        name = "cases") %>%   # counts col is named "cases" 
  mutate(                     # convert NA outcome to "Unknown"
    outcome = fct_na_value_to_level(
      outcome,
      level = "Unknown")) %>% 
  mutate(outcome = factor(outcome, levels = c("Recover", "Death"))) %>% 
  pivot_wider(                # pivot outcomes into separate columns
    values_from = cases,      
    names_from = outcome,
    names_expand = TRUE) %>%  # ensure Death appears in table even if none
  mutate(Death = replace_na(Death, 0)) %>% # If no Deaths, write as 0 (not NA)
  mutate(Known = Death + Recover) %>% # create number with known outcome
  arrange(-Known) %>%                 # order by total cases, descending
  adorn_totals(where = "row") %>%     # add a total row
  mutate(                             # add new columns
    pct_death = percent(Death / Known, 0.1),         # pct who died (1 decimal)
    pct_recover = percent(Recover / Known, 0.1)) %>% # pct who recovered 
  select(hospital, Known, Recover, pct_recover, Death, pct_death) # adjust order
```

```{r}
# convert hospital table to an HTML output

# define border styling
border_style <- officer::fp_border(color="black", width=1)

# convert hospital table to an HTML output
hosp_table_flx <- hosp_table %>% 
  
  qflextable() %>%  # convert to flextable
  
  add_header_row(   # add top header row
    top = TRUE,
    values = c("Hospital", "Known", "Recovered", "", "Died", "")) %>% 
  
  set_header_labels(
    Recover = "Count",
    pct_recover = "Percent",
    Death = "Count",
    pct_death = "Percent"
  ) %>% 
  
  # merge cells
  merge_at(i = 1:2, j = 1, part = "header") %>% # merge rows 1 & 2 in column 1 
  merge_at(i = 1:2, j = 2, part = "header") %>% # merge rows 1 & 2 in column 2
  merge_at(i = 1, j = 3:4, part = "header") %>% # merge columns 3 & 4 in header row 1
  merge_at(i = 1, j = 5:6, part = "header") %>% # merge columns 5 & 6 in header row 2  

  # add vertical lines to separate Recovered and Died sections
  vline(part = "all", j = 2, border = border_style) %>%   # at column 2 
  vline(part = "all", j = 4, border = border_style) %>%   # at column 4
  hline(part = "body", i = nrow(hosp_table)-1, border = border_style) %>%   # at last row

  # align number columns to center
  align_nottext_col("center") %>% 
  
  # Bold headers and total row
  bold(i = 1:2, bold = TRUE, part = "header") %>% 
  bold(i = nrow(hosp_table), bold = TRUE, part = "body") %>% 
  
  # background colors
  bg(part = "body", bg = "gray95") %>% 
  bg(j = nrow(hosp_table), i = ~pct_death >= 60, part = "body", bg = "red") %>% 
  
  #add footer
  add_footer_lines(
    values = str_glue(
      "Cases reported as of {max_report_date}. Hospitals with >=60% case fatality highlighted.",
      max_report_date = max(surv$date_report, na.rm = T))) %>% 
  italic(part = "footer") %>% 
  
  # add title
  add_header_lines(values = "Case outcomes by hospital") %>% 
  align(part = "header", i = 1, align = "center")


# print
hosp_table_flx
```


# Epi curve for `r params$district`

```{r}
# Generate weekly date sequence from Monday prior to first case 
weekly_breaks <- seq.Date(
      from = floor_date(min(surv$date_onset, na.rm=T), "week", week_start = 1), 
      to   = ceiling_date(max(surv$date_onset, na.rm=T), "week", week_start = 1), 
      by   = "week")     

# Plot epi curve
ggplot(data = surv) +              # set data
  geom_histogram(
    mapping = aes(x = date_onset), # date of onset
    breaks = weekly_breaks,        # weekly breaks defined above
    closed = "left",               # correct binning
    color = "darkblue",            # bar lines color
    fill = "lightblue")+           # bar fill color
  scale_x_date(                    # x-axis adjustments
    expand            = c(0,0),    # remove space below bars        
    date_breaks       = "4 weeks", # major gridlines        
    date_minor_breaks = "week",    # minor gridlines         
    labels = scales::label_date_short())+ # auto-efficent date labels
  scale_y_continuous(              # y-axis adjustments
    expand = c(0,0))+              # remove extra space by axis
  theme_minimal()+                 # white background
  theme(                           # other theme adjustments
    plot.caption = element_text(   # caption
      hjust = 0,                   # left align
      face = "italic"),            # italic
    axis.title = element_text(     # axis adjustments
      face = "bold"))+             # make bold
  labs(                            # plot labels
    title    = "Weekly case incidence ",
    subtitle = "",
    x        = "Week of symptom onset",
    y        = "Weekly cases reported",
    caption  = stringr::str_glue(
      "n = {n_cases} from {params$district}; Shown in Monday weeks by symptom onset.\nOnset dates range from {min_date} to {max_date}\n{n_missing} cases missing date of onset and not shown",
      n_cases = nrow(surv),
      min_date = format(min(surv$date_onset, na.rm=T), format = '%a %d %b %Y'),
      max_date = format(max(surv$date_onset, na.rm=T), format = '%a %d %b %Y'),
      n_missing = fmt_count(surv, is.na(date_onset)))
  )
```



